### 2018-05-29 Physics Notes/Paper 2/Paper 2.tex
Paper 2 notes update

@@

\paragraph{Derive $Fd=Q\Delta V$}
$$F=EQ \quad E=\frac{V}{d} \quad\therefore F=\frac{QV}{d} \quad\therefore Fd=QV$$
\paragraph{Particle in a uniform electric field}
\begin{center}
\includegraphics[width=8cm]{particle.jpg}
\end{center}
\end{document}

>>>>

\usepackage{float}

@@

\paragraph{Derive $Fd=Q\Delta V$}
$$F=EQ \quad E=\frac{V}{d} \quad\therefore F=\frac{QV}{d} \quad\therefore Fd=QV$$
\paragraph{Particle in a uniform electric field}
\includegraphics[width=8cm]{particle.jpg}
\subsubsection{Electric potential}
\textbf{Absolute electric potential} - At a point in an electric field, the work done per unit charge on a small positively charged object to move it from infinity to that point in the field
\subsection{Capacitance}
\subsubsection{Capacitance, parallel plate capacitor and energy stored in a capacitor}
\textbf{Capacitance} - The charge stored per unit p.d. of a capacitor\\
In the presence of an electric field, the polar molecules in the electric field will rotate, causing it to become polarised, and so a better dielectric. This is why polar molecules are used for dielectrics instead of non polar molecules
\subsubsection{Capacitor charge and discharge}
\includegraphics[width=6cm]{charge.png}
Voltage and charge follow the same pattern when charging and discharging, whereas current follows the opposite pattern.\\
The time constant(RC) is the time it until 37\% of the charge on the capacitor is remaining.
$$T_{\frac{1}{2}}=0.69RC$$
\subsection{Magnetic fields}
\subsubsection{Magnetic flux density}
\textbf{Magnetic flux density} - The magnetic force per unit length per unit current on a current carrying conductor at right angles to the field lines (Derive from F=BIl)
\subsubsection{Moving charges in a magnetic field}
$$F=BQv \quad \textrm{When the field is perpendicular to the velocity}$$
\paragraph{Cyclotron}
\includegraphics[width=10cm]{cyclotron.jpg}
This accelerates the particles to very high speeds, for use in medicine
\subsubsection{Magnetic flux and flux linkage}
$$\textrm{Magnetic flux:}\phi=BA \quad \textrm{Where B is normal to A}$$
\textbf{Flux} - $\phi=BA$ for a uniform magnetic field of flux density B that is perpendicular to an area A. Unit of Weber(Wb)\\
\textbf{Flux Linkage} - Through a coil of N turns, $N\phi=BAN\cos\theta$ where B is the magnetic flux density perpendicular to area A, unit also of Wb
\subsubsection{Electromagnetic induction}
\textbf{Faraday's law} - The induced emf in a circuit is equal to the rate of change of the magnetic flux linkage through the circuit\\
\textbf{Lenz's law} - When a current is induced in electromagnetic induction, the direction of the induced current is always such as to oppose the change that causes the current
\subsubsection{Alternating currents}
\textbf{Peak to peak} - The difference in the values between the two peak and the trough of an ac wave\\
\textbf{Root mean square} - The DC voltage that will give the same effect as the AC voltage
\paragraph{Oscilloscope}
$x$ axis - Time base\\
$y$ axis - Y sensitivity\\
The position and size of the waveform should be adjusted so that it takes up as much of the screen as possible, reducing uncertainty.
\subsubsection{Transformers}
\textbf{Eddy currents} - Induced currents in metal parts of ac machines\\
Inefficiencies:
\item Eddy currents form in the metal core of the transformer, leading to a loss in power
\item Resistance in the windings around the transformer resulting in energy wasted as heat
\item Magnetisation and demagnetisation of the core causes loss in energy
\end{itemize}
Power is transmitted at high voltage in transmission lines so they can be at a low resistance, and as power dissipated is $I^2R$, the lower the current, the less power is wasted
\section{Nuclear physics}
\subsection{Radioactivity}
\subsubsection{Rutherford scattering}
\includegraphics[width=8cm]{scatter.png}
Rutherford's experiment involved firing a beam of alpha particles at gold foil and measuring the paths of particles from the foil.
\item Gold was used as it was expected to have a large nucleus
\item The screen fluoresces when collided with
\item This showed the atom was mostly empty space with a positive nucleus
\end{itemize}
\begin{figure}[H]
    \centering
    \begin{minipage}{0.45\textwidth}
        \centering

\includegraphics[width=6cm]{foil.jpg}  
    \end{minipage}\hfill
    \begin{minipage}{0.45\textwidth}
        \centering
\begin{tabularx}{\textwidth}{|X|X|}
\hline
Observation&Explanation\\
\hline
Most electrons pass all the way through&Atoms are mostly empty space\\
\hline
Some are deflected&The atom has a positive centre\\
\hline
Some are deflected by significant angles&The positive charge is condensed in a small area\\
\hline
\end{tabularx}
    \end{minipage}
\end{figure}
\subsubsection{$\alpha$, $\beta$ and $\gamma$ radiation}
\begin{tabularx}{\textwidth}{|c|X|X|X|}
\hline
&Alpha&Beta&Gamma\\
\hline
Nature&2 Protons+2 Neutrons&High speed electron or positron&High energy photon\\
\hline
Range&Up to 10cm&Up to 1m&Infinite\\
\hline 
Deflection in a magnetic field&Deflected&Opposite direction to $\alpha$ particles and more easily deflected&Not deflected\\
\hline
Absorption&Paper&Aluminium&Lead \\
\hline
Ionisation&$10^4$ ions per mm&100 ions per mm&Very weak ionising effect\\
\hline
Energy of each particle&Constant for a given source&Varies up to a maximum for a given source&Constant for a given source\\
\hline
\end{tabularx}
\paragraph{Inverse square law for gamma radiation}
$$I=\frac{k}{x^2}$$
\paragraph{Safety}$ $\\
Those working in areas of high radiation, for example nuclear power plants, should wear film badges which monitor a persons exposure to radiation
\paragraph{Background radiation}$ $\\
Background radiation has many sources:
\begin{itemize}
\item Air (Radon gas)
\item Medical
\item Ground and buildings
\item Food and drink
\end{itemize}
\paragraph{Uses of radiation in medicine}$ $\\
The health benefits from using radiation in medicine almost always outweigh the risks, but doctors must protect themselves as repeated exposure is dangerous
\end{document}
